---
title: "R Notebook: Coaching Cafe: Using R for EDA"
output: html_notebook
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/Users/thomasmcneill/Documents/data/coaching_cafe')
```

load packages and set working directory

```{r}
# Clear environment
rm(list = ls())
# Clear console
cat("\014")

library(tidyr) ## spread
library(dplyr) ## mutate, filter, select, distinct, group_by, summarize
library(data.table) ## fread, pattern, merge, rbind, setorder
library(ggplot2) ## for plots
library(cowplot) ## print plots side by side


#setwd("/Users/thomasmcneill/Documents/data/coaching_cafe")
```
Note: there is a package to hit databases directly like SQL, but we don't use it at STU. I use R for standard data sets that have already been pulled and to merge those with external data sets (from CBP, USCIS, EOIR, etc)


Read in our data sets: 
1. Vehicle listings within 50 miles of my zip code 
2. Ratings for different models
```{r}
vehicle_listings <- fread("vehicle_listings.csv", check.names = T)
vehicle_ratings <- fread("vehicle_ratings.csv", check.names = T)
```

We can read them in one at a time (like above)
However, if we have a multiple data sets, it's easier to read them in all at once
```{r}
# create a list of the names of all files in the working directory, call the list "files"
files <- list.files(pattern = "*csv")

# read in all the files in the working directory, name them according to their filenames
for(i in 1:length(files)){
  assign(files[i],fread(files[i], check.names = TRUE))
}
```

List all columns in the listings df
```{r}
glimpse(vehicle_listings.csv)
```

```{r}
glimpse(vehicle_ratings.csv)
```

the pattern of the "car" column from ratings can be used as an identifier for each model/year
create the same column for the listings df so we can join the ratings to the listing df
```{r}
vehicle_listings.csv$car <- paste0(
  tolower(gsub(" ","-", vehicle_listings.csv$brand)),
  "/",
  tolower(gsub(" ","-", vehicle_listings.csv$model)),
  "/",
  tolower(gsub(" ","-", vehicle_listings.csv$productionYear)),
  "/")
```

Now we can join add ratings to the listings df
```{r}
listings <- merge(vehicle_listings.csv, vehicle_ratings.csv, by = "car", all.x = TRUE)
```

Clean up environment
```{r}
rm(list=ls(pattern = 'vehicle*'))
```

Confirm the join worked
```{r}
glimpse(listings)
```

Odometer was read in as a character; change it to numeric (remove commas first)
```{r}
listings$Odometer <- as.numeric(gsub(",","", listings$Odometer))
```

We can also view the df in a new tab
```{r}
View(listings)
```

Now that we have our final dataset, we can perform some EDA

First, let's create pivot tables
```{r}
car_year <- listings %>%
  group_by(productionYear) %>%        # rows (break out by production year)
  summarise(num_listings=n()) %>%     # count (total number or listings)
  setorder(-productionYear)           # order (production year descending)
car_year
```

We can also add column breakouts for pivot tables
```{r}
car_year_type <- listings %>%
  group_by(productionYear, bodyType) %>%        # rows (break out by production year)
  summarise(num_listings=n()) %>%     # count (total number or listings)
  spread(bodyType, num_listings, fill=0) %>%                # 
  setorder(-productionYear)           # order (production year descending)
car_year_type
```

And view in a separate tab
```{r}
View(car_year_type)
# or you can just click it in the console
```

Now let's create a few more pivot tables to use for plotting
```{r}
# filter to cars produced 2010 or later for clearer plots

# total number of listings by manufacturer
car_count <- listings %>%
  filter(productionYear>2009) %>%
  group_by(manufacturer) %>%
  summarise(num_listings=n()) %>%
  filter(num_listings>2) %>%
  setorder(-num_listings)

# average price of cars by manufacturer
car_price <- listings %>%
  filter(price>0) %>%
  filter(productionYear>2009) %>%
  group_by(manufacturer) %>%
  summarise(avg_price=mean(price)) %>%    # We can also use average as the pivot count instead of total listings
  setorder(-avg_price)

# average expert rating by manufacturer
car_exp_rating <- listings %>%
  filter(productionYear>2009) %>%
  filter(!is.na(expert_rating)) %>%
  group_by(manufacturer) %>%
  summarise(avg_rating=mean(expert_rating)) %>%
  setorder(-avg_rating)

# average consumer rating by manufacturer
car_cons_rating <- listings %>%
  filter(productionYear>2009) %>%
  filter(!is.na(consumer_rating)) %>%
  group_by(manufacturer) %>%
  summarise(avg_rating=mean(consumer_rating)) %>%
  setorder(-avg_rating)

# total number of listing by body type
car_type <- listings %>%
  filter(!is.na(bodyType)) %>%
  filter(bodyType != "Unavailable ") %>%
  group_by(bodyType) %>%
  summarise(num_listings=n()) %>%
  setorder(-num_listings)

car_price
```

Now let's create some plots with ggplot2
```{r}
##### Distribution of prices (excluding prices of 0)
# we can get a better view of the dist if we exclude the outliers to the right
listings %>%
  filter(price>0) %>%
  filter(price<100000) %>%
  ggplot(aes(price/1000)) +
  geom_histogram(bins=100)
```
The distribution skews right

Total number of listings and average prices by manufacturer
using pivots we've already created
```{r}
### manufacturer by number of listings
plot_count <- car_count %>%
  ggplot(aes(x=reorder(manufacturer, num_listings))) +
  geom_col(aes(y=num_listings)) +
  coord_flip() + 
  ggtitle("Number of Listings") +
  xlab("Manufacturer") + ylab("Number of Listings")

### manufacturer by average price
plot_price <- car_price %>%
  ggplot(aes(x=reorder(manufacturer, avg_price))) +
  geom_col(aes(y=avg_price)) + 
  coord_flip() + 
  ggtitle("Average Price") +
  xlab("Manufacturer") + ylab("Average Price")

# this function lets you show plots next to each other
plot_grid(plot_count, plot_price)
```
Average ratings by manufacturer: Expert v consumer ratings
```{r}
### manufacturer by average expert rating
plot_exp_rating <- car_exp_rating %>%
  ggplot(aes(x=reorder(manufacturer, avg_rating))) +
  geom_col(aes(y=avg_rating)) + 
  coord_flip() + 
  ggtitle("Expert Rating") +
  xlab("Manufacturer") + ylab("Avg Expert Rating")

plot_cons_rating <- car_cons_rating %>%
  ggplot(aes(x=reorder(manufacturer, avg_rating))) +
  geom_col(aes(y=avg_rating)) + 
  coord_flip() + 
  ggtitle("Consumer Rating") +
  xlab("Manufacturer") + ylab("Avg Consumer Rating")

plot_grid(plot_exp_rating, plot_cons_rating)
```
Ratings don't vary a lot

```{r}
### listings by body type
plot_type <- car_type %>%
  ggplot(aes(x=reorder(bodyType, num_listings))) +
  geom_col(aes(y=num_listings)) + 
  coord_flip() + 
  ggtitle("Body Type") +
  xlab("Body Type") + ylab("Number of Listings")

### listings by production year
plot_year <- car_year %>%
  filter(productionYear>1999) %>%
  ggplot(aes(x=productionYear)) +
  geom_col(aes(y=num_listings)) + 
  coord_flip() + 
  ggtitle("Production Year") +
  xlab("Production Year") + ylab("Number of Listings")

plot_grid(plot_type, plot_year)

```
Filter down the data set to meet your specifications
```{r}
##### Custom Search
custom_search <- listings %>%
  filter(price>0) %>%
  filter(price<15000) %>%
  filter(productionYear>2010) %>%
  filter(expert_rating>4.2)%>%
  filter(Odometer<110000) %>%
  setorder(manufacturer, model, price, Odometer) %>%
  filter(bodyType %in% c("Sedan","Hatchback","Sport Utility")) %>%
#  filter(manufacturer %in% c("Honda","Toyota","Subaru", "MAZDA")) %>%
  select(manufacturer, model, productionYear, bodyType, price, Odometer, color, expert_rating, 
         consumer_rating, itemCondition, driveWheelConfiguration, vehicleInteriorColor, 
         fuelType, fuelEfficiency, vehicleEngine, vehicleTransmission) %>%
  setorder(-expert_rating, price)
View(custom_search)
```

Create excel report 
```{r}

```

